\Large{Research Seminar Software Science} \\ \huge{Simulation of monadic effects in Haskell}

### 2018-06-02 paper.tex
addressing the problem with lower_r parser

    The parse \textbf{parseTwoItems} manages to parse the first two characters of the input string and returns the two parsed characters and remaining part of the input string which is not yet parsed. \\ \\
    In practice, the \textbf{parseTwoItems} is using two \textit{bind} infix operators. In the first \textit{bind}, the \textbf{oneItem} parser is eating the first character of the input string and passing the corresponding value to the function \textbf{\textbackslash x $\rightarrow$ oneItem}, which differently from the previous \textbf{parseOneItem} parser is returning another \textbf{oneItem} parser instead of the primitive \textbf{result} parser. In the second \textit{bind}, the  other \textbf{oneItem} parser is devouring the first character of the unparsed input string and passing that value as an input to the function \textbf{\textbackslash y $\rightarrow$ result(x, y)}. Finally, the resulting output will be handled by the \textbf{result} parser which will return the values of \textit{x} and \textit{y} being correspondingly the first and the second character of initial input string, together with the unparsed input string as a tuple. \\ \\ 
    The following example will better illustrate the outcome when parsing a string:
       \begin{center}
            \textbf{ghci$>$} parseTwoItems \textcolor{red}{``rera"} \\
            [((\textcolor{green}{`r'}, \textcolor{green}{`e'}), \textcolor{red}{``ra"})]
        \end{center}

@@

    \begin{tcolorbox}
    \begin{lstlisting}[language=Haskell]
      (+++) :: Parse a -> Parse a -> Parse a
      p +++ q  = \input -> (p input ++ q input)
    \end{lstlisting}
    \end{tcolorbox}

    \noindent At the moment, even though the first parser will succeed in parsing the input, the second parser will always be applied to the same input. This is not the desired behaviour from alternation of parsers. The goal is to allow the second parser to parse the input string only if the initial parser has failed in doing so. Additionally, if the first parser succeeds in parsing the input, the program is terminated and the parsed result is returned. This approach will introduce a deterministic behavior in the alternation of parsers. How can this be achieved? In order to achieve this attitude between the initial parser and the latter parser, the primitive \textit{zero} parser will be exploited. Recalling from section 5.2, the \textit{zero} parser will take any input and return absolutely nothing.

>>>>

    The parse \textbf{parseTwoItems} manages to parse the first two characters of the input string and returns the two parsed characters and remaining part of the input string which is not yet parsed. \\ \\
    In practice, the \textbf{parseTwoItems} is using two \textit{bind} infix operators. In the first \textit{bind}, the \textbf{oneItem} parser is eating the first character of the input string and passing the corresponding value to the function \textbf{\textbackslash x $\rightarrow$ oneItem}, which differently from the previous \textbf{parseOneItem} parser is returning another \textbf{oneItem} parser instead of the primitive \textbf{result} parser. In the second \textit{bind}, the  other \textbf{oneItem} parser is devouring the first character of the unparsed input string and passing that value as an input to the function \textbf{\textbackslash y $\rightarrow$ result(x, y)}. Finally, the resulting output will be handled by the \textbf{result} parser which will return the values of \textit{x} and \textit{y} being correspondingly the first and the second character of initial input string, together with the unparsed input string as a tuple. \\ \\ 
    The following example will better illustrate the outcome when parsing a string:
       \begin{center}
            \textbf{ghci$>$} parseTwoItems \textcolor{red}{``rera"} \\
            $\big[$((\textcolor{green}{`r'}, \textcolor{green}{`e'}), \textcolor{red}{``ra"})$\big]$
        \end{center}

@@

    \begin{tcolorbox}
    \begin{lstlisting}[language=Haskell]
      (+++) :: Parse a -> Parse a -> Parse a
      p +++ q  = \input -> (p input ++ q input)
    \end{lstlisting}
    \end{tcolorbox}
    \noindent The nature of the plus parser can be illustrated by defining the \textbf{islower\_r} parser which alternated two of the already defined parsers from the previous subsection: \textbf{islowercase} and \textbf{singleChar}.

    \begin{tcolorbox}
    \begin{lstlisting}[language=Haskell] 
        islower_r :: Parse Char
        islower_r  = islowercase +++ singleChar `r'
    \end{lstlisting}
    \end{tcolorbox}

    \noindent Now, let's apply the \textbf{islower\_r} parser to an input string that contains `r' as it's first character.

    \begin{center}
            \textbf{ghci$>$} islower\_r \textcolor{red}{``rera"} \\
            $\big[$(\textcolor{green}{`r'}, \textcolor{red}{``era"}), (\textcolor{green}{`r'}, \textcolor{red}{``era"})$\big]$
        \end{center}

     \noindent At the moment, even though the first parser will succeed in parsing the input, the second parser will always be applied to the same input.This is not the desired behaviour from alternation of parsers. The goal is to allow the second parser to parse the input string only if the initial parser has failed in doing so. Additionally, if the first parser succeeds in parsing the input, the program is terminated and the parsed result is returned. This approach will introduce a deterministic behavior in the alternation of parsers. How can this be achieved?
